Miniproject - Algebraic Geometry

### 2021-12-10 Algebraic geometry/submission.tex
added proof of important subspace lemma to alg geo

\begin{lemma}
    \label{prop:linear_transform_extpow}
    Let $A = (a_{ij}) \in \GL_d(k)$ and $v_1, ..., v_d \in V$.
    Then
    \begin{equation*}
        \Bigl( \sum_j a_{1j}v_j \Bigl) \wedge ... \wedge \Bigl( \sum_{j} a_{dj} v_j \Bigr) = \det(A) (v_1 \wedge ... \wedge v_d)
    \end{equation*}
\end{lemma}
\begin{proof}
    By a direct computation using Lemma~\ref{prop:basic_properties_exterior_product}, we find
    \begin{align*}
        &\Bigl( \sum_j a_{ij} v_j \Bigr) \wedge ... \wedge \Bigl( \sum_j a_{dj} v_j \Bigr) = \sum_{j_1, ..., j_d} a_{1j_1} ... a_{dj_d} (v_{j_1} \wedge ... \wedge v_{j_d}) \\
        = &\sum_{\pi \in S_d} a_{1\pi(1)} ... a_{d\pi(d)} (v_{\pi(1)} \wedge ... \wedge v_{\pi(d)}) \\
        = &\sum_{\pi \in S_d} a_{1\pi(1)} ... a_{d\pi(d)} \sgn(\pi) (v_1 \wedge ... \wedge v_d) \\
        = &(v_1 \wedge ... \wedge v_d) \ \sum_{\pi \in S_d} \sgn(\pi) \prod_{j = 1}^d a_{j\pi(j)} = \det(A) (v_1 \wedge ... \wedge v_d)
    \end{align*}
    where the last equality holds due to the Leibniz determinant formula.
\end{proof}
\begin{lemma}
    \label{prop:extpow_zero_iff_independent}
    For $v_1, ..., v_d \in V$ have
    \begin{equation*}
        v_1 \wedge ... \wedge v_d = 0 \ \Leftrightarrow \ v_1, ..., v_d \ \text{linearly dependent}
    \end{equation*}
\end{lemma}
\begin{proof}
    For the direction $\Leftarrow$, assume that $v_1, ..., v_d$ are not independent. 
    Then there is a nonzero vector $a_1 \in k^d$ with $\sum a_{1i} v_i = 0$.
    Clearly, we can extend $a_1$ to a basis $a_1, ..., a_d$ of $k^d$, which gives a matrix $A = (a_{ij}) \in \GL_d(k)$.

@@

    Direction $\Rightarrow$ TODO
\end{proof}
\begin{lemma}
    \label{prop:subspace_closed}
    Let $v \in V$ and $u \in \extpow^{d - 1} U$ for a linear subspace $U \leq V$. 
    If $v \wedge u \in \extpow^d U$ then $v \in U$ or $u = 0$.
\end{lemma}
\begin{proof}
    TODO
\end{proof}

>>>>

\begin{lemma}
    \label{prop:linear_transform_extpow}
    Let $A = (a_{ij}) \in \GL_d(k)$ and $v_1, ..., v_d \in V$.
    Then
    \begin{equation*}
        \Bigl( \sum_j a_{1j}v_j \Bigl) \wedge ... \wedge \Bigl( \sum_{j} a_{dj} v_j \Bigr) = \det(A) (v_1 \wedge ... \wedge v_d)
    \end{equation*}
\end{lemma}
\begin{proof}
    By a direct computation using Lemma~\ref{prop:basic_properties_exterior_product}, we find
    \begin{align*}
        &\Bigl( \sum_j a_{ij} v_j \Bigr) \wedge ... \wedge \Bigl( \sum_j a_{dj} v_j \Bigr) = \sum_{j_1, ..., j_d} a_{1j_1} ... a_{dj_d} (v_{j_1} \wedge ... \wedge v_{j_d}) \\
        = &\sum_{\pi \in S_d} a_{1\pi(1)} ... a_{d\pi(d)} (v_{\pi(1)} \wedge ... \wedge v_{\pi(d)}) \\
        = &\sum_{\pi \in S_d} a_{1\pi(1)} ... a_{d\pi(d)} \sgn(\pi) (v_1 \wedge ... \wedge v_d) \\
        = &(v_1 \wedge ... \wedge v_d) \ \sum_{\pi \in S_d} \sgn(\pi) \prod_{j = 1}^d a_{j\pi(j)} = \det(A) (v_1 \wedge ... \wedge v_d)
    \end{align*}
    where the last equality holds due to the Leibniz determinant formula.
\end{proof}
\begin{lemma}
    \label{prop:subspace_closed}
    Let $v \in V$ and $u \in \extpow^{d - 1} U$ for a linear subspace $U \leq V$. 
    If $v \wedge u \in \extpow^d U$ then $v \in U$ or $u = 0$.
\end{lemma}
\begin{proof}
    wlog $U$ is finite-dimensional, so there is a basis $u_1, ..., u_m$ of $U$.
    Then
    \begin{equation*}
        u = \sum_{\{i_1, ..., i_{d - 1}\} \in \{1, ..., m\}^{(d - 1)}} \lambda_{i_1, ..., i_{d - 1}} (u_{i_1} \wedge ... \wedge u_{i_{d - 1}})
    \end{equation*}
    So
    \begin{equation*}
        v \wedge u = \sum_{\{i_1, ..., i_{d - 1}\} \in \{1, ..., m\}^{(d - 1)}} \lambda_{i_1, ..., i_{d - 1}} (v \wedge u_{i_1} \wedge ... \wedge u_{i_{d - 1}}) \in \extpow^d U
    \end{equation*}
    It follows that
    \begin{align*}
        w := &\sum_{\{i_1, ..., i_{d - 1}\} \in \{1, ..., m\}^{(d - 1)}} \lambda_{i_1, ..., i_{d - 1}} (v \otimes u_{i_1} \otimes ... \otimes u_{i_{d - 1}}) \\
        &\in U^{\otimes d} + \sum_{i = 1}^{d - 1} V^{\otimes (i - 1)} \otimes \vspan\bigl\{ v \otimes v' + v' \otimes v \ \bigm| \ v, v' \in V \bigr\} \otimes V^{\otimes (d - i - 1)}
    \end{align*}
    hence $w$ is contained in
    \begin{equation*}
        \bigl( V \otimes U^{\otimes(d - 1)}\bigr) \cap \Bigl( U^{\otimes d} + \sum_{i = 1}^{d - 1} V^{\otimes (i - 1)} \otimes \vspan\bigl\{ v \otimes v' + v' \otimes v \bigr\} \otimes V^{\otimes (d - i - 1)} \Bigr)
    \end{equation*}
    However
    \begin{equation*}
        \bigl( V \otimes U^{\otimes(d - 1)}\bigr) \cap \Bigl( U^{\otimes d} + \sum_{i = 1}^{d - 1} V^{\otimes (i - 1)} \otimes \vspan\bigl\{ v \otimes v' + v' \otimes v \bigr\} \otimes V^{\otimes (d - i - 1)} \Bigr) = U^{\otimes d}
    \end{equation*}
    and so $w \in U^{\otimes d}$.

    Furthermore, we know by definition of $w$ that
    \begin{equation*}
        w \in \vspan\{v\} \otimes U^{\otimes(d - 1)}
    \end{equation*}
    However, if $v \notin U$ then $(\vspan\{v\} \otimes U^{\otimes(d - 1)}) \cap U^{\otimes d} = \{ 0 \}$ and we see that $w = 0$.
    We have that
    \begin{equation*}
        w = v \otimes \sum_{\{i_1, ..., i_{d - 1}\} \in \{1, ..., m\}^{(d - 1)}} \lambda_{i_1, ..., i_d} (u_{i_1} \otimes ... \otimes u_{i_{d - 1}})
    \end{equation*}
    As $v \neq 0$ (otherwise $v \in U$), we see that thus
    \begin{equation*}
        \sum_{\{i_1, ..., i_{d - 1}\} \in \{1, ..., m\}^{(d - 1)}} \lambda_{i_1, ..., i_d} (u_{i_1} \otimes ... \otimes u_{i_{d - 1}}) = 0
    \end{equation*}
    and so $u = 0$.
\end{proof}
\begin{lemma}
    \label{prop:extpow_zero_iff_independent}
    For $v_1, ..., v_d \in V$ have
    \begin{equation*}
        v_1 \wedge ... \wedge v_d = 0 \ \Leftrightarrow \ v_1, ..., v_d \ \text{linearly dependent}
    \end{equation*}
\end{lemma}
\begin{proof}
    For the direction $\Leftarrow$, assume that $v_1, ..., v_d$ are not independent. 
    Then there is a nonzero vector $a_1 \in k^d$ with $\sum a_{1i} v_i = 0$.
    Clearly, we can extend $a_1$ to a basis $a_1, ..., a_d$ of $k^d$, which gives a matrix $A = (a_{ij}) \in \GL_d(k)$.

@@

    For the other direction, let $v_1, ..., v_d$ be linearly independent.
    Assume that $v_1 \wedge ... \wedge v_d = 0$, i.e.
    \begin{equation*}
        v_1 \otimes ... \otimes v_d = \sum_{i = 0} u^{(i)}_1 \otimes ... \otimes u^{(i)}_d
    \end{equation*}
\end{proof}
